---
title: "Modeling"
author: "orlando"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(BasketballAnalyzeR) 
library(gridExtra)
library(tidymodels)
library(dplyr)
library(purrr)
library(ggplot2)
library(tidyverse)
library(corrplot)
library(modelr)
library(gapminder)
library(recipes)
```


```{r reading csv}

df <- read_csv("data/DFwithHourandLocation.csv")  
df$'Unnamed: 16' <- NULL
df$date <- as.Date(df$date)

# Trim spaces and replace with underscores in 'Name' column
df$Name <- gsub(" ", "_", trimws(df$Name))

#Removing rows where the MP =0 or MP = NA
df <- subset(df, !is.na(MP) & MP != 0)

df
```
Percebi que tenho um problema no DF original devido ao scrapping the identificado padrões idiomáticos diferentes, por exemplo temos como unique name 'nikola_joki?' e nikola_jokić que deveriam ser representados como um jogador e não como 2.


```{r identifying similar players names}
# Identify unique players in the main_df
unique_players <- unique(df$Name)

# Loop through each unique player
for (player in unique_players) {
  # Check if there are any similar player names
  similar_names <- agrep(player, unique_players, max.distance = 0.2, value = TRUE)
  
  # If there are similar names, correct them
  if (length(similar_names) > 1) {
    # Print the similar names for reference
    print(similar_names)
    
  }
}


```
Agora que já sabemos quais são os jogadores que estão dando problema no nome, podemos criar uma regra para corrigir eles de uma vez.
```{r fixing players names}

fix_player_names <- function(data) {
  name_mapping <- c(
    "mirza_teletović" = "mirza_teletovic",
    "mirza_teletovi?" = "mirza_teletovic",
    "boban_marjanović" = "boban_marjanovic",
    "boban_marjanovi?" = "boban_marjanovic",
    "bojan_bogdanović" = "bojan_bogdanovic",
    "bojan_bogdanovi?" = "bojan_bogdanovic",
    "goran_dragić" = "goran_dragic",
    "goran_dragi?" = "goran_dragic",
    "jusuf_nurkić" = "jusuf_nurkic",
    "jusuf_nurki?" = "jusuf_nurkic",
    "nikola_jokić" = "nikola_jokic",
    "nikola_joki?" = "nikola_jokic",
    "dario_šarić" = "dario_saric",
    "dario_šari?" = "dario_saric",
    "dario_교ri?" = "dario_saric",
    "miloš_teodosi?" = "milos_teodosic",
    "miloš_teodosić" = "milos_teodosic",
    "ante_žiži?" = "ante_zizic",
    "ante_žižić" = "ante_zizic",
    "bogdan_bogdanovi?" = "bogdan_bogdanovic",
    "bogdan_bogdanović" = "bogdan_bogdanovic",
    "nikola_mirotić" = "nikola_mirotic",
    "nikola_miroti?" = "nikola_mirotic",
    "donatas_motiej?nas" = "donatas_motiejunas",
    "donatas_motiejūnas" = "donatas_motiejunas",
    "ersan_i̇lyasova" = "ersan_ilyasova",
    "ersan_?lyasova" = "ersan_ilyasova", 
    "nikola_vučević" = "nikola_vucevic",
    "nikola_vu?evi?" = "nikola_vucevic",
    "jonas_valančiūnas" = "jonas_valanciunas",
    "jonas_valan?i?nas" = "jonas_valanciunas",
    "dāvis_bertāns" = "davis_bertans",
    "d?vis_bert?ns" = "davis_bertans",
    "dairis_bert?ns" = "davis_bertans",
    "kristaps_porziņģis" = "kristaps_porzingis",
    "kristaps_porzi??is" = "kristaps_porzingis",
    "dairis_bert?ns" = "dairis_bertans",
    "damjan_rudež" = "damjan_rudez",
    "sasha_vujačić" = "sasha_vujacic",
    "nikola_peković" = "nikola_pekovic",
    "tomáš_satoranský" = "tomas,satoransky",
    "luka_don?i?" = "luka_doncic",
    "džanan_musa" = "dzanan_musa",
    "vlatko_?an?ar" = "vlatko_cancar",
    "anžejs_pase??iks" = "anzejs_pasecnkis",
    "alen_smailagi?" = "alen_smailagic",
    "luka_šamani?" = "luka_samanic",
    "alperen_?engün" = "alperen_sengun"
  )
  
  data$Name <- ifelse(data$Name %in% names(name_mapping), name_mapping[data$Name], data$Name)
  
  return(data)
}

# Call the fix_player_names() function to fix the names in the "data" data frame
df <- fix_player_names(df)



```
Checking if some name was forgotten on the new df df

```{r}
# Identify unique players in the main_df
unique_players2 <- unique(df$Name)

# Loop through each unique player
for (player in unique_players2) {
  # Check if there are any similar player names
  similar_names2 <- agrep(player, unique_players2, max.distance = 0.05, value = TRUE)
  
  # If there are similar names, correct them
  if (length(similar_names2) > 1) {
    # Print the similar names for reference
    print(similar_names2)
    
  }
}
print('---')
#Doing an extra check to see if there are any players with "?" or the problematic characters in the name
# A list of some players were found and add on the fix_data function
for (player in unique_players2) {

  if (grepl("\\?|ć|ž|č|ū|ā|ģ|š|ü|교", player)) {
    # Print the player names with non-UTF-8 characters
    print(player)

  }
}
```


Overview of the data
```{r overview}
# Summary of the DataFrame
summary_df <- summary(df)
print(summary_df)
```
Data Cleaning:
Notei que as colunas númericas estão ficando com NA ao invés de 0 no df, abaixo vou preenche-las com 0 e investigar o caso onde AST% = -1000. Depois de investigar, eu acredito que esses valores negativos sejam algum tipo de bug na tabela original, portanto a % será transformada em 0
Edit: Após conversa com o professor, decidimos remover, por enquanto, as colunas de % para evitar o possível conflito onde 0Sucessos/0Tentativas = 0% e 0Sucessos/VáriasTentativas = 0%.
```{r}

# Fill NA values with 0 in the specified columns
df$PF <- replace(df$PF, is.na(df$PF), 0)
df$'TS%' <- replace(df$'TS%', is.na(df$'TS%'), 0)
df$'eFG%' <- replace(df$'eFG%', is.na(df$'eFG%'), 0)
df$FTr <- replace(df$FTr, is.na(df$FTr), 0)
df$BPM <- replace(df$BPM, is.na(df$BPM), 0)
df$'3PAr' <- replace(df$'3PAr', is.na(df$'3PAr'), 0)
df$'AST%' <- replace(df$'AST%', is.na(df$'AST%'), 0)
df$'TOV%' <- replace(df$'TOV%', is.na(df$'TOV%'), 0)
df$'+/-' <- replace(df$'+/-', is.na(df$'+/-'),0)


# Replace negative values with 0 in the filtered DataFrame
df$'AST%'[df$'AST%' < 0] <- 0


# Remove specified columns of FG%, X3P% and FT%
df <- df %>%
  select(-c('FG%', '3P%', 'FT%'))


# Remove duplicate rows
df <- distinct(df)


summary2_df <- summary(df)
print(summary2_df)


```

Criando colunas de mês dentro da season
```{r}

# Extract the year and month from the date column
df <- df %>%
  mutate(year = lubridate::year(date),
         month = lubridate::month(date))

# Find the unique combinations of season and year to determine the season start months
season_start_dates <- df %>%
  distinct(season, year, month)

season_start_dates$season_month <- 1

# Initialize the counter for season and month
current_season <- season_start_dates$season[1] #2016
current_season_month <- 1

# Loop through each row in the dataframe
for (i in 1:nrow(season_start_dates)) {
  # Check if the season and month have changed
  if (season_start_dates$season[i] != current_season) {
    current_season_month <- 1
    current_season <- season_start_dates$season[i]
    current_season_month <- current_season_month + 1
  }
  else{
    season_start_dates$season_month[i] <- current_season_month
    current_season_month <- current_season_month + 1
  }
}

# Join season_start_dates with df
df <- df %>%
  left_join(season_start_dates, by = c("season", "year","month"))

df


```

Usarei a função nest() para agrupar o df por jogador
```{r nest}

# Assuming your dataframe is named df
nested_df <- df %>% 
  group_by(Name) %>% 
  nest(.key="data")



# Print the nested data
nested_df

```
Criando colunas extras no nest para total de jogos e minutos jogados para cada jogador
```{r}
# Define a function to calculate the sum of MP
sum_MP <- function(data) {
  sum(data$MP)
}

# Define a function to count the occurrences of the player
count <- function(data) {
  nrow(data)
}

nested_df <- nested_df %>%
  mutate(sum_MP = map_dbl(data, sum_MP),
         count = map_int(data, count))


# Print the nested data with sum_MP and count columns
nested_df

```


Testando como está o nested para os jogadores
```{r testing player}
# Filter the nested data for "joel_embiid"
player_data <- nested_df[nested_df$Name == "joel_embiid", ]$data

# Print the nested data for "joel_embiid"
print(player_data)

```
Criando Feature Engineering columns
```{r feature engineering}

# Function to calculate rest days
calculate_rest_days <- function(data) {
  data$rest_days <- c(0, diff(data$date))
  return(data)
}

# Apply the calculate_rest_days function to nested_df
nested_df <- nested_df %>%
  mutate(data = purrr::map(data, ~ calculate_rest_days(.)))

#Creating the columns next_opp and next_PTS 
nested_df <- nested_df %>%
  mutate(data = map(data, ~ {
    df <- .x
    df$next_opp <- c(df$opp[-1], NA)
    df$next_PTS <- c(df$PTS[-1], NA)
    df
  }))


# View the updated nested_df with the rest_days column within each nested dataset
nested_df
```
Conferindo se as mudanças deram certo
```{r}
player_name <- "joel_embiid"

# Filter the nested data for "joel_embiid"
joel_embiid_data <- nested_df %>% 
  filter(Name == player_name) %>%
  pull(data) %>%
  pluck(1)

# Print the nested data for "joel_embiid"
print(joel_embiid_data)
```

Opto por seguir o passo-a-passo do Tidy Modeling em primeiro momento utilizando todas as variavéis do $data dos jogadores.
```{r}

clone_nested <- nested_df %>%
    filter(count > 20)


lm_model <- function(data) {
  # First, remove factors with only one level from the predictors [Ex: When the player played in only 1 team]
  predictors <- select(data, where(~ is.factor(.) && n_distinct(.) > 1))
  
  #First we delete the last row of the data that will contain next_PTS = NAN
  data <- data %>% slice(-n())
  
  # Ensure that the `strata` column is present and has compatible sizes
  if (!"season_month" %in% colnames(data) || !is.numeric(data$season_month)) {
    return(NULL)
  }
  
  # Fit the linear regression model
  set.seed(502)
  data_split <- initial_split(data, prop = 0.80, strata = season_month)
  
  data_train <- training(data_split)
  data_test <- testing(data_split)
  
  lm_model <- linear_reg() %>% set_engine("lm")
  
  lm_wflow <-
    workflow() %>%
    add_model(lm_model) %>%
    add_variables(outcome = next_PTS, predictors = names(predictors))
  
  lm_fit <- fit(lm_wflow, data_train)
  lm_fit
}


clone_nested <- clone_nested %>%
  mutate(model = map(data,lm_model))

clone_nested


```
O lm_model acima será avaliado no futuro

Vou fazer a initial_split aqui e deixar os dados já divididos para facilitar as operações no futuro
```{r splitting into train and test data}

split_data <- function(data) {
  #First we delete the last row of the data that will contain next_PTS = NAN
  data <- data %>% slice(-n())
  
  set.seed(502)
  data_split <- initial_split(data, prop = 0.80, strata = season_month)
  data_train <- training(data_split)
  data_test <- testing(data_split)
  list(data_train = data_train, data_test = data_test)
}

clone_nested <- clone_nested %>%
  mutate(splitado = map(data,split_data))
clone_nested

```

```{r extracting data_train and data_test}


clone_nested <- clone_nested %>%
  mutate(data_train = map(splitado, pluck, "data_train"),
         data_test = map(splitado, pluck, "data_test"))
clone_nested


```
```{r removing the splitado column}
clone_nested <- clone_nested %>%
  select(-splitado)
clone_nested
```




```{r}

simple_lm <- function(data,data_train){

  recipe(next_PTS ~ ., data = data_train) %>%
    step_select(where(~ is.factor(.) && n_distinct(.) > 1)) %>%
    step_dummy(all_nominal_predictors(), -all_outcomes())
}

clone_nested <- clone_nested %>%
  mutate(simple = mapply(simple_lm, data, data_train, SIMPLIFY = FALSE))
clone_nested


```


```{r}

recipe_model <- function(recipe) {
  lm_model <- linear_reg() %>% set_engine("lm")
  
  lm_wflow <- 
    workflow() %>%
    add_model(lm_model) %>%
    add_recipe(recipe)
  lm_wflow
}

clone_nested <- clone_nested %>%
  mutate(recipe_wflow = map(simple, ~recipe_model(.x)))
clone_nested


```


```{r}
#Let’s estimate both the recipe and model using a simple call to fit():
recipe_fit <- function(wflow, data_train) {
  set.seed(502)
  data_train_df <- as.data.frame(data_train)  # Convert data_train to a data frame
  print(data_train_df)

  lm_fit <- fit(wflow, data = data_train_df)

  lm_fit
}

clone_nested <- clone_nested %>%
  mutate(rec_fit = mapply(recipe_fit, recipe_wflow, data_train, SIMPLIFY = FALSE))

clone_nested


```

Edit: Nesse ponto me veio a ideia de ter o data_train e o data_test como duas colunas externas ao nest() para não ficar chamando o data_split repetidas vezes. Voltei no código e fiz isso, vou remover em algum update futuro.
```{r}
clone_nested %>%
  filter(Name == "a.j._hammons") %>%
  pull(data_train)
```



```{r}

test <- clone_nested %>% 
  filter(Name == "joel_embiid") %>%
  pull(simple)

test

```


