---
title: "Analises"
author: "orlando"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BasketballAnalyzeR) 
library(gridExtra)
library(tidymodels)
library(dplyr)
library(purrr)
library(ggplot2)
library(tidyverse)
library(corrplot)
library(modelr)
library(gapminder)
library(recipes)

```


```{r reading csv}

df <- read_csv("data/cleanedData.csv")  
df
```


Univariate Analysis
```{r UA}
# Histogram of MP variable
ggplot(df, aes(x = MP)) +
  geom_histogram(binwidth = 5) +
  xlab("MP") +
  ylab("Frequency") +
  ggtitle("Distribution of Minutes Played")

# Histogram of PTS variable
ggplot(df, aes(x = PTS)) +
  geom_histogram(binwidth = 5) +
  xlab("Points") +
  ylab("Frequency") +
  ggtitle("Distribution of Points")

# Bar chart of a categorical variable
ggplot(df, aes(x = team)) +
  geom_bar() +
  xlab("Team") +
  ylab("Count") +
  ggtitle("Number of Games Played by Team")
```
Bivariate Analysis
```{r BA}

# Scatter plot of MP x PTS
ggplot(df, aes(x = MP, y = PTS)) +
  geom_point() +
  xlab("MP") +
  ylab("Points") +
  ggtitle("MP vs. Points")

# Correlation matrix of most important variables
correlation_matrix <- cor(df[c("MP","PTS", "AST", "TRB","STL","BLK","TOV","PF","FG","FGA","3P","3PA","FT","FTA")])
corrplot(correlation_matrix, method = "circle")

# Correlation matrix with the advanced stats column
correlation_matrix <- cor(df[c("PTS","AST", "TRB", "TS%","eFG%","3PAr","FTr","ORB%","DRB%","TRB%","AST%","STL%","BLK%", "TOV%","USG%","ORtg","DRtg","BPM")])
corrplot(correlation_matrix, method = "circle")

numerical_df <- df %>% 
  mutate(team = as.numeric(as.factor(team)),
         opp = as.numeric(as.factor(opp)))

# Correlation matrix with teams and seasonal data
correlation_matrix <- cor(numerical_df[c("PTS","AST", "TRB", "opp","team","season")])
corrplot(correlation_matrix, method = "circle")

#correlation among everything


correlation_matrix <- cor(numerical_df[c("MP","PTS", "AST", "TRB","STL","BLK","TOV","PF","FG","FGA","3P","3PA","FT","FTA", "TS%","eFG%","3PAr","FTr","ORB%","DRB%","TRB%","AST%","STL%","BLK%", "TOV%","USG%","ORtg","DRtg","BPM","opp","team","season")])
corrplot(correlation_matrix, method = "circle")




```
Multivariate Analysis


```{r}
# Box plot of points by team
ggplot(df, aes(x = team, y = PTS)) +
  geom_boxplot() +
  xlab("Team") +
  ylab("Points") +
  ggtitle("Points by Team") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Line plot of points over time
df_season <- df %>% group_by(season) %>% summarize(avg_points = mean(PTS))
ggplot(df_season, aes(x = season, y = avg_points)) +
  geom_line() +
  xlab("Season") +
  ylab("Average Points") +
  ggtitle("Average Points per Season") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Usarei a função nest() para agrupar o df por jogador
```{r nest}

# Assuming your dataframe is named df
nested_data <- df %>% 
  group_by(Name) %>% 
  nest(.key="data")



# Print the nested data
nested_data

```
Criando colunas extras no nest para total de jogos e minutos jogados para cada jogador
```{r}
# Define a function to calculate the sum of MP
sum_MP <- function(data) {
  sum(data$MP)
}

# Define a function to count the occurrences of the player
count <- function(data) {
  nrow(data)
}

nested_data <- nested_data %>%
  mutate(sum_MP = map_dbl(data, sum_MP),
         count = map_int(data, count))


# Print the nested data with sum_MP and count columns
nested_data

```


Testando como está o nested para os jogadores
```{r testing player}
# Filter the nested data for "joel_embiid"
player_data <- nested_data[nested_data$Name == "joel_embiid", ]$data

# Print the nested data for "joel_embiid"
print(player_data)

```
Criando Feature Engineering columns
```{r feature engineering}

# Function to calculate rest days
calculate_rest_days <- function(df) {
  df$rest_days <- c(0, diff(df$date))
  return(df)
}

# Apply the calculate_rest_days function to nested_data
nested_data <- nested_data %>%
  mutate(data = purrr::map(data, ~ calculate_rest_days(.)))

#Creating the columns next_opp and next_PTS 
nested_data <- nested_data %>%
  mutate(data = map(data, ~ {
    df <- .x
    df$next_opp <- c(df$opp[-1], NA)
    df$next_PTS <- c(df$PTS[-1], NA)
    df
  }))


# View the updated nested_data with the rest_days column within each nested dataset
nested_data
```
Conferindo se as mudanças deram certo
```{r}
player_name <- "joel_embiid"

# Filter the nested data for "joel_embiid"
joel_embiid_data <- nested_data %>% 
  filter(Name == player_name) %>%
  pull(data) %>%
  pluck(1)

# Print the nested data for "joel_embiid"
print(joel_embiid_data)
```

Opto por seguir o passo-a-passo do Tidy Modeling em primeiro momento utilizando todas as variavéis do $data dos jogadores.
```{r}

clone_nested <- nested_data %>%
    filter(count > 20)


lm_model <- function(data) {
  # First, remove factors with only one level from the predictors [Ex: When the player played in only 1 team]
  predictors <- select(data, where(~ is.factor(.) && n_distinct(.) > 1))
  
  #First we delete the last row of the data that will contain next_PTS = NAN
  data <- data %>% slice(-n())
  
  # Ensure that the `strata` column is present and has compatible sizes
  if (!"season_month" %in% colnames(data) || !is.numeric(data$season_month)) {
    return(NULL)
  }
  
  # Fit the linear regression model
  set.seed(502)
  data_split <- initial_split(data, prop = 0.80, strata = season_month)
  
  data_train <- training(data_split)
  data_test <- testing(data_split)
  
  lm_model <- linear_reg() %>% set_engine("lm")
  
  lm_wflow <-
    workflow() %>%
    add_model(lm_model) %>%
    add_variables(outcome = next_PTS, predictors = names(predictors))
  
  lm_fit <- fit(lm_wflow, data_train)
  lm_fit
}


clone_nested <- clone_nested %>%
  mutate(model = map(data,lm_model))

clone_nested




```
O lm_model acima será avaliado no futuro

Vou fazer a initial_split aqui e deixar os dados já divididos para facilitar as operações no futuro
```{r splitting into train and test data}

split_data <- function(data) {
  set.seed(502)
  data_split <- initial_split(data, prop = 0.80, strata = season_month)
  data_train <- training(data_split)
  data_test <- testing(data_split)
  list(data_train = data_train, data_test = data_test)
}

clone_nested <- clone_nested %>%
  mutate(splitado = map(data,split_data))
clone_nested

```

```{r extracting data_train and data_test}


clone_nested <- clone_nested %>%
  mutate(data_train = map(splitado, pluck, "data_train"),
         data_test = map(splitado, pluck, "data_test"))
clone_nested


```
```{r removing the splitado column}
clone_nested <- clone_nested %>%
  select(-splitado)
clone_nested
```




```{r}

simple_lm <- function(data,data_train){

 recipe_obj <- recipe(next_PTS ~ ., data = data_train) %>%
    step_select(where(~ is.factor(.) && n_distinct(.) > 1)) %>%
    step_slice(-n()) %>%
    step_dummy(all_nominal_predictors(), -all_outcomes())
  recipe_obj
  
}

clone_nested <- clone_nested %>%
  mutate(simple = map2(data,data_train, simple_lm))
clone_nested


```


```{r}

recipe_model <- function(recipe) {
  lm_model <- linear_reg() %>% set_engine("lm")
  
  lm_wflow <- 
    workflow() %>%
    add_model(lm_model) %>%
    add_recipe(recipe)
  lm_wflow
}

clone_nested <- clone_nested %>%
  mutate(recipe_wflow = map(simple, ~recipe_model(.x)))
clone_nested


```
Removida a parte onde rodava um modelo linear reg aqui, decido deixar isso para arquivos futuros. 


```{r}

test <- clone_nested %>% 
  filter(Name == "joel_embiid") %>%
  pull(simple)

test

```




Testando a lenda urbana de que o Jokic pontua mais em jogos no horário da tarde. O desempenho do atual MVP também será posto nesse primeiro gráfico para uma comparação visual. Tentando ver provas visuais de que o horário do jogo pode ser significante para alguns jogadores

```{r}
# Define the player names
player_names <- c("joel_embiid", "nikola_jokic")

# Filter the nested data for specific players
filtered_data <- nested_data %>%
  filter(Name %in% player_names) %>%
  mutate(data = map(data, ~ .x %>% unnest(cols = c(PTS, hour_float))))

# Combine the filtered and unnested data into a single dataframe
combined_data <- bind_rows(filtered_data$data, .id = "Name") %>%
  mutate(Name = recode(Name, "1" = "nikola_jokic", "2" = "joel_embiid")) %>%
  group_by(Name, hour_float, PTS) %>%
  mutate(count = n())

# Print the combined data
print(combined_data)

# Plot the graph
ggplot(combined_data, aes(x = hour_float, y = PTS, color = Name, size = count)) +
  geom_point() +
  labs(x = "Hour", y = "PTS", color = "Player", size = "Frequência") +
  scale_color_manual(values = c("joel_embiid" = "red", "nikola_jokic" = "blue"))



```

```{r correlating hour of the game and pts}



# Define a function to perform the statistical analysis for each nested group
hour_model <- function(data) {
  # Fit the linear regression model
  lm(PTS ~ hour_float, data = data)
}

models <- map(nested_data$data, hour_model)

#Cloning the nested_data for this battery of tests
clone_nested2 <- nested_data %>%
  mutate(hourmodel = map(data,hour_model))
clone_nested2


```
```{r adding residuals to the nest}

clone_nested2 <- clone_nested2 %>%
  mutate(
    resids = map2(data,hourmodel,add_residuals)
  )
clone_nested2
```
```{r unnesting}
resids <- unnest(clone_nested2, resids)
resids

```
```{r plotting}


resids %>%
 ggplot(aes(hour_float, resid)) +
 geom_line(aes(group = Name), alpha = 1 / 3) +
 geom_smooth(se = FALSE)
#> `geom_smooth()` using method = 'gam'


```
```{r faceting by team}
resids %>%
 ggplot(aes(hour_float, resid, group = Name)) +
 geom_line(alpha = 1 / 3) +
 facet_wrap(~team)

```
```{r glance}

glance <- clone_nested2 %>%
  mutate(glance = map(hourmodel,broom::glance)) %>%
  unnest(glance, .drop = TRUE)
glance


```

```{r}
glance %>%
 arrange(r.squared)

```
```{r}

good_fit <- filter(glance, abs(r.squared) > 0.75 )

good_fit

unnested_good_fit <- good_fit %>%
  unnest(cols = c(data))

unnested_good_fit

name_counts <- unnested_good_fit %>%
  group_by(Name, count) %>%
  summarise() %>%
  ungroup() %>%
  arrange(desc(count))


name_counts


```

Baseado no gráfico, que não mostrou nenhum padrão de rendimento baseado no horário e que as pessoas com melhores rendimentos em horários especifico são pessoas com poucos jogos, decidi não realizar testes ANOVA nem me aprofundar nesse caminho.

