---
title: "04 - Watching the past"
author: "orlando"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(zoo)
library(readr)


```

Initially, the cleaned CSV should be read and the unnecessary columns should be removed before proceeding with further analysis.Date will be no longer useful, neither year-month-day columns. We'll keep only season and season_month for time columns.
Location is not relevant in our models.
Since plus-minus statistics can have limitations and may not always accurately reflect a player's individual performance since they can be influenced by various factors, including the performance of teammates and opponents. They will also be removed from the model.

Furthermore, the columns related to "Advanced Box Score Stats," which are estimations based on team possessions or plays, will be eliminated. This includes columns such as TS%, eFG%, 3PAr, FTr, ORB%, DRB%, TRB%, AST%, STL%, BLK%, TOV%, USG%, ORtg, DRtg, and BPM.

```{r reading csv}

df <- read_csv("data/cleanedData.csv") 

calculate_rest_days <- function(data) {
  data <- data %>% 
    group_by(Name) %>%
    mutate(rest_days = c(0, diff(date))) %>%
    mutate(rest_days = ifelse(rest_days > 30, 0, rest_days))
  
  
  return(data)
}

df <- calculate_rest_days(df)

columns_to_be_removed <- c("+/-", "TS%", "eFG%", "3PAr", "FTr", "ORB%", "DRB%", "TRB%", "AST%", "STL%", "BLK%", "TOV%", "USG%", "ORtg", "DRtg", "BPM", "date", "location", "year", "month", "day")

df <- df[, !(names(df) %in% columns_to_be_removed)]

# Convert "opp" column to factor for grouping
df$opp <- as.factor(df$opp)



df
```



Function to calculate the rolling median for a given column

This function calculates the rolling median for a specified column in a data frame.
It creates additional columns for the last n games' median of the player's performance
and the last n games' median of the player's performance against the specific opponent.

Arguments:
- data: The data frame containing the player's game data.
- column: The name of the column for which the rolling median will be calculated.
- n: The number of games to consider in the rolling window.
Returns:
 The data frame with additional columns for the rolling medians and n/2 rolling medians against the opponent..

Example usage:
 df <- calculate_rolling_median(data, "FG", 10)
 Calculates the rolling median of Field Goals (FG) for the last 10 games.
 Edit: The function now excludes the actual game and return the median for the 10 games before the actual game. [More realistic approach]

The resulting data frame will have additional columns with names following the format "LN_stat" for the player's last 10 game medians and "LN_stat_opp" for the player's last N game medians against the specified team.

To use the function, pass the data frame, column, and the desired length of medians.

```{r}

# Create a function to calculate the median
calculate_median <- function(x) {
  if (length(x) == 1) {
    return(0)
  } else {
    return(median(x[-length(x)], na.rm = TRUE))
  }
}


# Function to calculate the rolling median for a given column
calculate_rolling_median <- function(data, column, n) {
  data <- data %>% group_by(Name)
  
  column_name <- paste0("L", n, "_", column)
  
  # Apply the function to calculate the rolling median using a window of size 10
  data <- data %>%
    mutate("{column_name}" := rollapply(.data[[column]], width = (n+1), FUN = calculate_median,
                                  fill = NA, align = "right", partial = TRUE)) %>%
    ungroup()
  
  # Group the data by player name and opponent
  data <- data %>% group_by(Name, opp)
  
  n_opp <- floor(n/2)
  
  column_name_opp <- paste0("L", n_opp, "_", column,"_opp")

  # Apply the function to calculate the median using rolling window of size 10 per opponent
  data <- data %>%
    mutate("{column_name_opp}" := rollapply(.data[[column]], width = (n_opp+1), FUN = calculate_median, fill = NA, align = "right",partial=TRUE))

  data <- data %>% ungroup()
  
  
  return(data)
}


```
Applying the function on the desired columns and n=10

```{r} 
columns_to_watch <- c("FG", "FGA", "3P", "3PA", "FT", "FTA", "ORB", "DRB", "TRB", "AST", "STL", "BLK", "TOV", "PF", "PTS")

for (column in columns_to_watch) {
  print(column)
  df <- calculate_rolling_median(df, column, 10)
}

```

Creating featured engineering for the column Points per Minute (ppm) and applying the changes on it. 


```{r}

df$ppm <- df$PTS / df$MP

df <- calculate_rolling_median(df, "ppm", 10)


```


Now removing the single game data, the same as columns_to_watch and keeping the df forward the past games stats plus the basic data of the actual game in it:
```{r}
columns_to_remove <- c("FG", "FGA", "3P", "3PA", "FT", "FTA", "ORB", "DRB", "TRB", "AST", "STL", "BLK", "TOV", "PF", "ppm")

df <- df[, -which(names(df) %in% columns_to_remove)]


```



Saving the "Watching the Past" df


```{r}

df
write.csv(df, file = "data/wtp.csv", row.names = FALSE)



```




