---
title: "Modeling"
author: "orlando"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gridExtra)
library(tidymodels)
library(dplyr)
library(purrr)
library(ggplot2)
library(tidyverse)
library(corrplot)
library(modelr)
library(gapminder)
library(recipes)
```


```{r reading csv}

df <- read_csv("data/cleanedData.csv")  

df
```


Overview of the data
```{r overview}
# Summary of the DataFrame
summary_df <- summary(df)
print(summary_df)
```

Usarei a função nest() para agrupar o df por jogador
```{r nest}

# Assuming your dataframe is named df
nested_df <- df %>% 
  group_by(Name) %>% 
  nest(.key="data")

# Print the nested data
nested_df

```

Criando colunas extras no nest para total de jogos e minutos jogados para cada jogador
```{r}
# Define a function to calculate the sum of MP
sum_MP <- function(data) {
  sum(data$MP)
}

# Define a function to count the occurrences of the player
count <- function(data) {
  nrow(data)
}

nested_df <- nested_df %>%
  mutate(sum_MP = map_dbl(data, sum_MP),
         count = map_int(data, count))


# Print the nested data with sum_MP and count columns
nested_df

```


Testando como está o nested para os jogadores
```{r testing player}
# Filter the nested data for "joel_embiid"
player_data <- nested_df[nested_df$Name == "joel_embiid", ]$data

# Print the nested data for "joel_embiid"
print(player_data)

```
Criando Feature Engineering columns
```{r feature engineering}

# Function to calculate rest days
calculate_rest_days <- function(data) {
  data$rest_days <- c(0, diff(data$date))
  return(data)
}

# Apply the calculate_rest_days function to nested_df
nested_df <- nested_df %>%
  mutate(data = purrr::map(data, ~ calculate_rest_days(.)))

#Creating the columns next_opp and next_PTS 
nested_df <- nested_df %>%
  mutate(data = map(data, ~ {
    df <- .x
    df$next_opp <- c(df$opp[-1], NA)
    df$next_PTS <- c(df$PTS[-1], NA)
    df
  }))


# View the updated nested_df with the rest_days column within each nested dataset
nested_df
```
Conferindo se as mudanças deram certo
```{r}
player_name <- "joel_embiid"

# Filter the nested data for "joel_embiid"
joel_embiid_data <- nested_df %>% 
  filter(Name == player_name) %>%
  pull(data) %>%
  pluck(1)

# Print the nested data for "joel_embiid"
print(joel_embiid_data)
```

Opto por seguir o passo-a-passo do Tidy Modeling em primeiro momento utilizando todas as variavéis do $data dos jogadores.
```{r}

clone_nested <- nested_df %>%
    filter(count > 20)


lm_model <- function(data) {
  # First, remove factors with only one level from the predictors [Ex: When the player played in only 1 team]
  predictors <- select(data, where(~ is.factor(.) && n_distinct(.) > 1))
  
  #First we delete the last row of the data that will contain next_PTS = NAN
  data <- data %>% slice(-n())
  
  # Ensure that the `strata` column is present and has compatible sizes
  if (!"season_month" %in% colnames(data) || !is.numeric(data$season_month)) {
    return(NULL)
  }
  
  # Fit the linear regression model
  set.seed(502)
  data_split <- initial_split(data, prop = 0.80, strata = season_month)
  
  data_train <- training(data_split)
  data_test <- testing(data_split)
  
  lm_model <- linear_reg() %>% set_engine("lm")
  
  lm_wflow <-
    workflow() %>%
    add_model(lm_model) %>%
    add_variables(outcome = next_PTS, predictors = names(predictors))
  
  lm_fit <- fit(lm_wflow, data_train)
  lm_fit
}


clone_nested <- clone_nested %>%
  mutate(model = map(data,lm_model))

clone_nested


```
O lm_model acima será avaliado no futuro

Vou fazer a initial_split aqui e deixar os dados já divididos para facilitar as operações no futuro
```{r splitting into train and test data}

split_data <- function(data) {
  #First we delete the last row of the data that will contain next_PTS = NAN
  data <- data %>% slice(-n())
  
  set.seed(502)
  data_split <- initial_split(data, prop = 0.80, strata = season_month)
  data_train <- training(data_split)
  data_test <- testing(data_split)
  list(data_train = data_train, data_test = data_test)
}

clone_nested <- clone_nested %>%
  mutate(splitado = map(data,split_data))
clone_nested

```

```{r extracting data_train and data_test}


clone_nested <- clone_nested %>%
  mutate(data_train = map(splitado, pluck, "data_train"),
         data_test = map(splitado, pluck, "data_test"))
clone_nested


```
```{r removing the splitado column}
clone_nested <- clone_nested %>%
  select(-splitado)
clone_nested
```




```{r}

simple_lm <- function(data,data_train){

  recipe(next_PTS ~ ., data = data_train) %>%
    step_select(where(~ is.factor(.) && n_distinct(.) > 1)) %>%
    step_dummy(all_nominal_predictors(), -all_outcomes())
}

clone_nested <- clone_nested %>%
  mutate(simple = mapply(simple_lm, data, data_train, SIMPLIFY = FALSE))
clone_nested


```


```{r}

recipe_model <- function(recipe) {
  lm_model <- linear_reg() %>% set_engine("lm")
  
  lm_wflow <- 
    workflow() %>%
    add_model(lm_model) %>%
    add_recipe(recipe)
  lm_wflow
}

clone_nested <- clone_nested %>%
  mutate(recipe_wflow = map(simple, ~recipe_model(.x)))
clone_nested


```


```{r}
#Let’s estimate both the recipe and model using a simple call to fit():
recipe_fit <- function(wflow, data_train) {
  set.seed(502)
  data_train_df <- as.data.frame(data_train)  # Convert data_train to a data frame
  print(data_train_df)

  lm_fit <- fit(wflow, data = data_train_df)

  lm_fit
}

clone_nested <- clone_nested %>%
  mutate(rec_fit = mapply(recipe_fit, recipe_wflow, data_train, SIMPLIFY = FALSE))

clone_nested


```

Edit: Nesse ponto me veio a ideia de ter o data_train e o data_test como duas colunas externas ao nest() para não ficar chamando o data_split repetidas vezes. Voltei no código e fiz isso, vou remover em algum update futuro.
```{r}
clone_nested %>%
  filter(Name == "a.j._hammons") %>%
  pull(data_train)
```



```{r}

test <- clone_nested %>% 
  filter(Name == "joel_embiid") %>%
  pull(simple)

test

```


Essa ideia foi deixada em stand by até descobrir qual o melhor modelo para esse tipo de dados
